cardinalR: Generating interesting high-dimensional data structures

### 2025-06-24 paper-cardinalR.Rmd
Updated cube section

The initial points are drawn from a uniform cube centered at the origin: $X_1, X_2, X_3 \sim U(-0.5, 0.5)$. The center of the cube is $(0, 0, 0)$. A spherical hole of radius $0.5$ is carved out by filtering out points whose Euclidean distance from the origin is less than or equal to $0.5$: Remove if $\sqrt{X_1^2 + X_2^2 + X_3^2} \leq 0.5$. This results in a shell-like point cloud that fills the cube but avoids the spherical center. For $p > 3$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $3\text{-}D$ structure into $p\text{-}D$, where $j = 4, \dots, p$.

@@

The curve is constructed by drawing uniformly spaced inputs and applying a second-degree polynomial transformation. Let $X_1 \sim U(\text{range}[1], \text{range}[2])$ be the independent variable. A raw polynomial basis of degree 2 is used to form $X_2 = X_1 - X_1^2 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.

This creates a smooth parabolic arc that opens downward, with jitter in the vertical direction to simulate measurement noise or intrinsic variability.

For $p > 2$, the dataset includes Gaussian noise in the remaining dimensions via `gen_noisedims()`:

$$
X_j \sim N(0, 0.1^2), \quad j = 3, \dots, p.
$$

@@

The shape is generated using a third-degree raw polynomial basis expansion. Let $X_1 \sim U(\text{range}[1], \text{range}[2])$ be the base input. A cubic transformation with vertical jitter is used to define $X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.

When $p > 2$, the remaining dimensions consist of independent Gaussian noise features generated via `gen_noisedims()`:

$$
X_j \sim N(0, 0.1^2)\$ for \$j = 3, \dots, p.
$$

>>>>

The initial points are drawn from a uniform cube centered at the origin: $X_1, X_2, X_3 \sim U(-0.5, 0.5)$. The center of the cube is $(0, 0, 0)$. A spherical hole of radius $0.5$ is carved out by filtering out points whose Euclidean distance from the origin is less than or equal to $0.5$: remove if $\sqrt{X_1^2 + X_2^2 + X_3^2} \leq 0.5$. This results in a shell-like point cloud that fills the cube but avoids the spherical center. For $p > 3$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $3\text{-}D$ structure into $p\text{-}D$, where $j = 4, \dots, p$.

@@

The curve is constructed by drawing uniformly spaced inputs and applying a second-degree polynomial transformation. Let $X_1 \sim U(\text{range}[1], \text{range}[2])$ be the independent variable. A raw polynomial basis of degree 2 is used to form $X_2 = X_1 - X_1^2 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$. This creates a smooth parabolic arc that opens downward, with jitter in the vertical direction to simulate noise. For $p > 2$, Gaussian noise $X_j \sim N(0, 0.1^2)$ is added to embed the $2\text{-}D$ structure into $p\text{-}D$, where $j = 3, \dots, p$.

@@

The shape is generated using a third-degree raw polynomial basis expansion. Let $X_1 \sim U(\text{range}[1], \text{range}[2])$ be the base input. A cubic transformation with vertical jitter is used to define $X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$. For $p > 2$, Gaussian noise $X_j \sim N(0, 0.1^2)$ is added to embed the $2\text{-}D$ structure into $p\text{-}D$, where $j = 3, \dots, p$.

### 2025-06-24 paper-cardinalR.tex
Updated cube section

The initial points are drawn from a uniform cube centered at the origin: \(X_1, X_2, X_3 \sim U(-0.5, 0.5)\). The center of the cube is \((0, 0, 0)\). A spherical hole of radius \(0.5\) is carved out by filtering out points whose Euclidean distance from the origin is less than or equal to \(0.5\): Remove if \(\sqrt{X_1^2 + X_2^2 + X_3^2} \leq 0.5\). This results in a shell-like point cloud that fills the cube but avoids the spherical center. For \(p > 3\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(3\text{-}D\) structure into \(p\text{-}D\), where \(j = 4, \dots, p\).

@@

The curve is constructed by drawing uniformly spaced inputs and applying a second-degree polynomial transformation. Let \(X_1 \sim U(\text{range}[1], \text{range}[2])\) be the independent variable. A raw polynomial basis of degree 2 is used to form \(X_2 = X_1 - X_1^2 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).

This creates a smooth parabolic arc that opens downward, with jitter in the vertical direction to simulate measurement noise or intrinsic variability.

For \(p > 2\), the dataset includes Gaussian noise in the remaining dimensions via \texttt{gen\_noisedims()}:

\[
X_j \sim N(0, 0.1^2), \quad j = 3, \dots, p.
\]

@@

The shape is generated using a third-degree raw polynomial basis expansion. Let \(X_1 \sim U(\text{range}[1], \text{range}[2])\) be the base input. A cubic transformation with vertical jitter is used to define \(X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).

When \(p > 2\), the remaining dimensions consist of independent Gaussian noise features generated via \texttt{gen\_noisedims()}:

\[
X_j \sim N(0, 0.1^2)\$ for \$j = 3, \dots, p.
\]

>>>>

The initial points are drawn from a uniform cube centered at the origin: \(X_1, X_2, X_3 \sim U(-0.5, 0.5)\). The center of the cube is \((0, 0, 0)\). A spherical hole of radius \(0.5\) is carved out by filtering out points whose Euclidean distance from the origin is less than or equal to \(0.5\): remove if \(\sqrt{X_1^2 + X_2^2 + X_3^2} \leq 0.5\). This results in a shell-like point cloud that fills the cube but avoids the spherical center. For \(p > 3\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(3\text{-}D\) structure into \(p\text{-}D\), where \(j = 4, \dots, p\).

@@

The curve is constructed by drawing uniformly spaced inputs and applying a second-degree polynomial transformation. Let \(X_1 \sim U(\text{range}[1], \text{range}[2])\) be the independent variable. A raw polynomial basis of degree 2 is used to form \(X_2 = X_1 - X_1^2 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\). This creates a smooth parabolic arc that opens downward, with jitter in the vertical direction to simulate noise. For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.1^2)\) is added to embed the \(2\text{-}D\) structure into \(p\text{-}D\), where \(j = 3, \dots, p\).

@@

The shape is generated using a third-degree raw polynomial basis expansion. Let \(X_1 \sim U(\text{range}[1], \text{range}[2])\) be the base input. A cubic transformation with vertical jitter is used to define \(X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\). For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.1^2)\) is added to embed the \(2\text{-}D\) structure into \(p\text{-}D\), where \(j = 3, \dots, p\).
